---
title: "Foreign students data analysis"
author: "Ian Hodgson"
date: "2025-07-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

## Share of foreign students

``` {r, echo = F}
foreign <- read_csv("data/IPEDS/foreign_enrollment.csv") %>% 
  clean_names()

foreign %>% filter(is.na(rmfrgncn_drvef2023)) %>% nrow()
```

More than half of observations do not list the number of foreign-born students. Will need to look into the reporting requirements for this indicator. In the meantime, what can we say about the type of institutions that don't report this figure.

``` {r, echo = F}

foreign %>% mutate(foreign_na_23 = is.na(rmfrgncn_drvef2023), 
                   foreign_na_18 = is.na(rmfrgncn_drvef2018_rv)) %>% 
  group_by(c21enprf_hd2023, 
           foreign_na_23, 
           foreign_na_18) %>% 
  summarise(count = n())

```

The type of institution (mostly 2-year, mostly 4-year, mostly graduate, etc.) doesn't tell us much. It's also concerning that some schools will report the figure one year and not the next.

### Control of school

``` {r, echo = F} 
foreign %>% mutate(foreign_na_23 = is.na(rmfrgncn_drvef2023), 
                   foreign_na_18 = is.na(rmfrgncn_drvef2018_rv)) %>% 
  group_by(control_hd2023, 
           foreign_na_23, 
           foreign_na_18) %>% 
  summarise(count = n())

```
Control of school (public, private, non-profit) doesn't seem to shed any light either. 

I'm going to leave this issue for now. I just need to find examples of schools with high or growing foreign shares and with declining or flat enrollment.


``` {r, echo = f}
foreign_top20 <- foreign %>% filter(!is.na(rmfrgncp_drvef2018_rv)) %>% 
  arrange(control_hd2023, 
          desc(rmfrgncp_drvef2023)) %>% 
  group_by(control_hd2023) %>% 
  mutate(count = 1:n()) %>% 
  filter(count < 21) %>% 
  select(institution_name, stabbr_hd2023, rmfrgncp_drvef2023)
```

Some notable schools on the list. Among the the publics: 
* UC-Davis and UC-Irvine
* Univeristy of Illinois Urbana-Champaign

Among the non-profits: 
* RISDI
* The New School
* NYU
* Boston University
* Brandeis
* Carnegie Mellon
* Emory
* Grinnell

## Enrollment declines

``` {r, echo = F}
enrollment <- read_csv("data/IPEDS/fall_enrollment.csv") %>% clean_names()

enrollment_long <- enrollment %>% 
  select(-x24) %>% 
  pivot_longer(cols = c(3, 5:23)) %>% 
  mutate(year = str_extract(name, "[0-9]{4}") %>% as.numeric())
```

Let's look at enrollment for the 20 non-profit schools with the highest shares of foreign students.

``` {r, echo = F}
foreign_top20 %>% 
  filter(control_hd2023 == 2) %>% 
  head(n = 3) %>% 
  left_join(enrollment_long) %>% 
  rename(enrollment = value) %>% 
  mutate(chart_name = paste(institution_name, ",", rmfrgncp_drvef2023, "per cent foreign (2023)")) %>% 
  ggplot() +
  geom_line(aes(x = year, y = enrollment)) + 
  theme_minimal() + 
  facet_wrap(facets = "chart_name", ncol = 1, scales = "free_y")
```